Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-16 paper.qmd
Updated Figure 17

right_top_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Reds 2") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

@@

middle_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Greens 3") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

@@

right_bottom_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Purples 2") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

@@

imge_error_sample_outside <- pixels_gathered_outside |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

@@

imge_error_sample_inside <- pixels_gathered_inside |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

```{r}
#| label: fig-model-error-mnist
#| fig-cap: "The $10\\text{-}D$ model error in layout a of the MNIST digit 1 dataset show a pattern. Most low model errors are distributed along the big non-linear clutser, while most large model errors are distributed along the small cluster. Along the non-linear cluster, the angle of digit 1 changes. Some images have large errors due to their deviation from the non-linear cluster, which makes them anomalies. The images associated with large model errors shows different patterns of digit 1."
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5
#| out-width: 80%

highd_error_2d_mnist + (imge_error_sample_outside/ imge_error_sample_inside / right_top_img / middle_img / right_bottom_img)
```

>>>>

right_top_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  ggtitle("(c)") +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Reds 2") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.31))

@@

middle_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  ggtitle("(d)") +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Greens 3") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.31))

@@

right_bottom_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  ggtitle("(e)") +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Purples 2") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.31))

@@

imge_error_sample_outside <- pixels_gathered_outside |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  ggtitle("(a)") +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.31))

@@

imge_error_sample_inside <- pixels_gathered_inside |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  ggtitle("(b)") +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.31))

```{r}
#| label: fig-model-error-mnist
#| fig-cap: "The $10\\text{-}D$ model error in \\gD{} layout of the MNIST digit 1 data shows a pattern. Most low model errors are distributed along the big non-linear cluster, while most large model errors are distributed along the small cluster. The images associated with large model errors shows different patterns of digit 1, some inside (b) the non-linear structure and others outside (a). Along the non-linear cluster, the angle of digit 1 changes (c-e)."
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5
#| out-width: 80%

highd_error_2d_mnist + (imge_error_sample_outside / imge_error_sample_inside / right_top_img / middle_img / right_bottom_img) 
```
